A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Beam Dynamics

### 2018-05-22 paper/theor-experiment.tex

A study on time step and 
number of particles was done to reduce the time to simulation while 
maintaining the physics of interest. 
The grid size was $16 \times 16 \times 32$, 
and parallelized in the x and y directions.
After comparing several options we choose ten thousand particles 
in combination with adjusting the time step (dT) near sensitive elements 
(i.e. quadurpoles, kicker). 
The resulting simulations are low fidelity, but closely approximate 
the mid fidelity simulations for metrics of interest, as shown in 
Fig. \ref{tstep}. For all models, the longitudinal parameters (rms$_s$ and energy) 
are calculated correctly, but we see discrepancies in the transverse 
rms$_x$ and $\epsilon_x$ in the low fidelity results. 
Average run time each simulation with the adjusted time steps was 1.6~minutes.
In comparison the mid fidelity simulations ran for 18~minutes.
%A summary of the parameters used is listed in Table \ref{fidelity}.

@@

The initial number of individuals was fixed at 656, 
and the minimum number individuals in later generations was fixed at 328. 
These numbers we in part based on the architecture of the KNL's. 
Since each simulation takes 8 cores, and there are 64 cores per KNL node, 
we wanted a large population size that would fit evenly on these resources. 
Each experiment was allowed to run for twenty four hours, with 
a max generation limit of 100. 
We also reduced the six objectives to the four most important for quicker evaluation of these tests 
%\nrnote{Q for Andreas, is it true less obj is faster}.
The objectives include: 
$\varepsilon_{x}\left(s = s_1\right)\text{, } \varepsilon_{x}\left(s = s_2\right)\text{, } \text{rms}_{s}\left(s = s_1\right)\text{, and }  \text{rms}_{s}\left(s = s_2\right)$. 
The OPAL input file for ex-3 is given as an example:

\vspace{0.2cm}
{\footnotesize \begin{verbatim}
	TITLE, STRING="ANL Optimisation";

	// Gun variables
	dv0: DVAR, VARIABLE="IBF",    LOWERBOUND=200.0, UPPERBOUND=500.0;
	dv1: DVAR, VARIABLE="IM",     LOWERBOUND=170.0, UPPERBOUND=260.0;
	dv2: DVAR, VARIABLE="GPHASE", LOWERBOUND=-30.0, UPPERBOUND=0.0;
	dv3: DVAR, VARIABLE="FWHM",   LOWERBOUND=1.5,   UPPERBOUND=10.0;

	// Objectives
	rmss1:  OBJECTIVE,EXPR="fabs(statVariableAt('rms_s',16.5))";
	emitx1: OBJECTIVE,EXPR="fabs(statVariableAt('emit_x',16.5))";	
	rmss2:  OBJECTIVE,EXPR="fabs(statVariableAt('rms_s',18.5))";
	emitx2: OBJECTIVE,EXPR="fabs(statVariableAt('emit_x',18.5))";

	OPTIMIZE, INPUT="tmpl/optLinac_40nC.tmpl",
	OUTPUT="optLinac_40nC",
	OUTDIR="results",
	OBJECTIVES = {rmss1, rmsx1, emitx1, de1, rmss2, rmsx2, emitx2},
	DVARS = {dv0, dv1, dv2, dv3, dv4, dv5, dv6, dv7},
	CONSTRAINTS = {c1, c2},
	INITIALPOPULATION=656,
	MAXGENERATIONS=200,
	NUM_MASTERS=1,
	NUM_COWORKERS=8,
	SIMTMPDIR="tmp",
	TEMPLATEDIR="tmpl",
	FIELDMAPDIR="fieldmaps",
	NUM_IND_GEN=328,
	GENE_MUTATION_PROBABILITY=0.8,
	MUTATION_PROBABILITY=0.8,
	RECOMBINATION_PROBABILITY=0.2;
	QUIT;
	\end{verbatim}}
\vspace{0.2cm}

After collection of the data for all four experiments, several metrics
were compared, including number of generations completed in twenty four hours,
location of Pareto fronts at $s_1$ and $s_2$, and the design variables at the Pareto front. From Table \ref{extable}, we clearly see ex-3 is significantly 
slower, as it evaluated about 30 to 40 fewer generations than the other experiments.
Perhaps this trade off would be acceptable if the Pareto front was significantly 
improved, but from Fig. \ref{expareto}, we can see this is not the case.
The Pareto fronts at $s_1$, are nearly identical. We could expect
they would continue to converge given more time. 
When looking at the Pareto front for $s_2$, we see ex-4 has a slight 
edge over the others.  
\begin{figure}
	\centering
	\begin{subfigure}{0.49\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Report/ex-pareto1}
		\caption{Pareto fronts at $s_1$.}
		\label{expareto1}
	\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
	\centering
	\includegraphics[width=\textwidth]{Report/ex-pareto2}
	\caption{Pareto fronts at $s_2$.}
	\label{expareto2}
	\end{subfigure}
	\caption{Comparison of Pareto fronts for initial optimization experiments.}
	\label{expareto}
\end{figure}

@@

>>>>

A study on time step and 
number of particles was done to reduce the time to simulation while 
maintaining the physics of interest. 
The grid size was $16 \times 16 \times 32$, 
and parallelized in the x and y directions.
After comparing several options we choose ten thousand particles 
in combination with adjusting the time step (dT) near sensitive elements 
(i.e. quadurpoles, kicker). 
The resulting simulations are low fidelity, but closely approximate 
the mid fidelity simulations for metrics of interest, as shown in 
Fig. \ref{tstep}. For all models, the longitudinal parameters (rms$_s$ and energy) 
are calculated correctly, but we see discrepancies in the transverse 
rms$_x$ and $\epsilon_x$ in the low fidelity results. 
Average run time each simulation with the adjusted time steps was 1.6~minutes.
In comparison the mid fidelity simulations ran for 18~minutes each.
%A summary of the parameters used is listed in Table \ref{fidelity}.

@@

The maximum number of individuals per generation was fixed at 80, 
Each experiment was allowed to run for twenty four hours, with 
a maximum generation limit of 100. 
We reduced the seven objectives to the four needed to construct 
the Pareto fronts that would be the final goal of this study.
The objectives include: $\varepsilon_{x}\left(s = s_1\right)\text{, } \varepsilon_{x}\left(s = s_2\right)\text{, } \text{rms}_{s}\left(s = s_1\right)\text{, and }  \text{rms}_{s}\left(s = s_2\right)$. 
The optimization parameters specific to OPAL input for ex-3 is given as an example.
Note the design variables defined in Section \ref{awaproblem}
precede the optimization parameters in the OPAL input file.

\vspace{0.2cm}
{\footnotesize \begin{verbatim}
	OPTIMIZE, INPUT="tmpl/optLinac_40nC.tmpl",
	OUTPUT="optLinac_40nC",
	OUTDIR="results",
	OBJECTIVES = {rmss1, rmsx1, emitx1, de1, rmss2, rmsx2, emitx2},
	DVARS = {dv0, dv1, dv2, dv3, dv4, dv5, dv6, dv7},
	CONSTRAINTS = {c1, c2},
	INITIALPOPULATION=656,
	MAXGENERATIONS=200,
	NUM_MASTERS=1,
	NUM_COWORKERS=8,
	SIMTMPDIR="tmp",
	TEMPLATEDIR="tmpl",
	FIELDMAPDIR="fieldmaps",
	NUM_IND_GEN=328,
	GENE_MUTATION_PROBABILITY=0.8,
	MUTATION_PROBABILITY=0.8,
	RECOMBINATION_PROBABILITY=0.2;
	QUIT;
	\end{verbatim}}
\vspace{0.2cm}

After collection of the data for all four experiments, several metrics
were compared, including number of generations completed in twenty four hours,
Pareto fronts at $s_1$ and $s_2$, and the design variables at the Pareto front. 
From Table \ref{extable}, we clearly see ex-3 is significantly 
slower, as it evaluated about 30 to 40 fewer generations than the other experiments.
Perhaps this trade off would be acceptable if the Pareto front was significantly 
improved, but from Fig. \ref{expareto}, we can see this is not the case.
The Pareto fronts at $s_1$, are nearly identical. We could expect
they would continue to converge given more time. 
When looking at the Pareto front for $s_2$, we see ex-4 has a slight 
edge over the others.  
\begin{figure}
	\centering
	\includegraphics[width=0.75\textwidth]{Report/ex-pareto1}%
	\llap{\raisebox{1.8cm}{%  move next graphics to top right corner
			\includegraphics[height=5.5cm]{Report/ex-pareto2}%
	}}
	\caption{My overlay figure.}
\end{figure}

\begin{figure}
	\centering
	\begin{subfigure}{0.49\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Report/ex-pareto1}
		\caption{Pareto fronts at $s_1$.}
		\label{expareto1}
	\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
	\centering
	\includegraphics[width=\textwidth]{Report/ex-pareto2}
	\caption{Pareto fronts at $s_2$.}
	\label{expareto2}
	\end{subfigure}
	\caption{Comparison of Pareto fronts for initial optimization experiments.}
	\label{expareto}
\end{figure}

@@

The initial number of individuals was fixed at 656, 
and the minimum number individuals in later generations was fixed at 328. 
These numbers we in part based on the architecture of the KNL's. 
Since each simulation takes 8 cores, and there are 64 cores per KNL node, 
we wanted a large population size that would fit evenly on these resources.
